\thetitle

### 2020-11-13 sun_phases/spin_qudit_tomo.tex
added some discussions in sections on reconstruction error, which are now 'complete' (though subject to future revision)

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reconstruction error bound}
\label{sec:bound}

For the practically minded, the mere existence of solutions to a problem is less interesting than the exposition of a particular solution.
On a high level, a spin qudit tomography protocol consists of
\begin{enumerate*}
\item selecting a set of measurement axes,
\item collecting measurement data on spin projection onto these axes, and then
  \label{tomo:measure}
\item processing the collected data to reconstruct the state of the spin qudit.
\end{enumerate*}
Whereas step \ref{tomo:measure} can involve a host of platform-dependent technical challenges, here we discuss the steps to take before and after collecting measurement data.

@@

A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\set{\Pi_{\v v\mu}}$ that will be measured in an experiment.
By a simple change of basis, measuring these projectors is equivalent to measuring the set of transition operators $\v T_V\equiv\set{T_{\v v\ell,0}}$.
By flattening the operators in $\v T_V$ into column vectors $\oket{T_{\v v\ell,0}}$, we can construct the {\it measurement matrix}
\begin{align}
  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell,0}} \bra{\v v\ell},
  \label{eq:meas_mat}
\end{align}
A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured transition operators in $\v T_V$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
In this case $M_V$ must be full rank, with $d^2$ nonzero singular values.
Indexing these singular values $M^V_k$ and the corresponding (normalized) right singular vectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots,x^V_{k,\abs{V}})$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal qudit operators
\begin{align}
  Q^V_k \equiv \sum_j q^V_{kj} T_j,
  &&
  \v q^V_k \equiv \f{\v x^V_k}{M^V_k},
\end{align}
where for shorthand we use a combined index $j=\p{\v v_j,\ell_j}$ to specify both a measurement axis $\v v_j$ and a degree $\ell_j$, which identify the transition operator $T_j\equiv T_{\v v_j\ell_j,0}$.
These operators allow us to expand any state $\rho$ of a $d$-level spin qudit in the form
\begin{align}
  \rho = \sum_{k=1}^{d^2} \rho_k^V \p{Q_k^V}^\dag,
  &&
  \rho_k^V \equiv \bk{Q^V_k}_\rho.
\end{align}
Given empirical estimates $\tilde T_j$ of the expectation values $\bk{T_j}_\rho$, an empirical estimate $\tilde\rho_V$ of $\rho$ is then
\begin{align}
  \tilde\rho_V \equiv \sum_k \tilde\rho^V_k \p{Q^V_k}^\dag,
  \label{eq:reconstructed_state}
\end{align}
where
\begin{align}
  \tilde\rho^V_k \equiv \sum_j q^V_{kj} \tilde T_j
  \approx \sum_j q^V_{kj} \bk{T_j}_\rho
  = \bk{Q^V_k}_\rho
  = \rho^V_k.
\end{align}
The measurement matrix $M_V$ allows us to make concrete statements about the statistical error between the empirical estimate $\tilde\rho_V$ and the true state $\rho$.
Assume, for example, that the estimates $\tilde T_j$ are equal to $\bk{T_j}_\rho$ up to uncorrelated noise with variance at most $\epsilon^2$:
\begin{align}
  \tilde T_j = \bk{T_j}_\rho + \epsilon_j,
  &&
  \bbk{\epsilon_j \epsilon_{j'}} \le \epsilon^2 \delta_{jj'},
  \label{eq:error_assumption}
\end{align}
where $\set{\epsilon_j}$ are independent random variables and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging.
In this case, the mean squared error with which $\tilde\rho^V_k$ approximates $\rho^V_k$ is
\begin{align}
  \Bbk{\abs{\tilde\rho^V_k-\rho^V_k}^2}
  &= \Bbk{\p{\tilde\rho^V_k-\rho^V_k}^*\p{\tilde\rho^V_k-\rho^V_k}} \\
  &= \sum_{j,j'} \p{q^V_{kj}}^* q^V_{kj'}\,
  \bbk{\epsilon_j \epsilon_{j'}} \\
  &\le \sum_j \abs{q^V_{kj}}^2 \epsilon^2
  = \p{\f{\epsilon}{M^V_k}}^2.
\end{align}
Using the fact that the operators $Q^V_k$ are orthonormal, we can therefore bound the mean squared (Euclidean) distance between $\tilde\rho_V$ and $\rho$ as
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  \le \epsilon^2 \S_V^2,
  \label{eq:bound_eps}
\end{align}
where $\norm{X} \equiv \sqrt{\obk{X|X}} = \sqrt{\tr\p{X^\dag X}}$ is the (Hilbert-Schmidt, or Frobenius) norm of $X$, and the {\it classical error scale}
\begin{align}
  \S_V^2 \equiv \sum_k\f1{\p{M^V_k}^2} = \norm{M_V^{-1}}^2,
\end{align}
is the squared norm of the right inverse $M_V^{-1}$ of $M_V$.
Note that the classical error scale $\S_V^2$ diverges if $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.

Computing the classical error scale $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of a transition operator $T_{\ell m}$ is preserved under rotations, which implies that the unitary
\begin{align}
  U \equiv \sum_{\ell,m} \ket{\ell m} \obra{T_{\ell m}},
\end{align}
block-diagonalizes the measurement matrix:
\begin{align}
  U M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
\end{align}
where the $\p{2\ell+1}\times\abs{V}$ blocks are
\begin{align}
  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell,0}} \bra{\v v}
  = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
\end{align}
As the singular values of $M_V$ are invariant under unitary transformations, it follows that
\begin{align}
  \S_V^2 = \sum_\ell \S_{V\ell}^2,
  &&
  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2,
\end{align}
where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time (see Appendix \ref{sec:compute} for details on efficiently computing $\S_{V\ell}^2$).
If we assume that $\abs{V}\sim d$, then computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).

@@

The assumption that observables can be estimated up to uncorrelated noise with maximal variance $\epsilon^2$, summarized by Eq.~\eqref{eq:error_assumption}, is reasonable when measurement error is dominated by experimental sources of noise.
However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:error}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$, with shot noise the dominant source of error.
In this case, the mean squared distance between the approximation $\tilde\rho_V$ and the true state $\rho$ can be compactly expressed as as an inner product of the {\it covariance matrices}
\begin{align}
  \C_\ell\sp{\rho} \equiv \sum_{m,n}
  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n}
\end{align}
where $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$, and the {\it noise matrices}
\begin{align}
  \N_{V\ell} \equiv M_{V\ell} \diag\sp{M_{V\ell}^{-1}
    \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag,
\end{align}
where $\diag\sp{X}$ sets all off-diagonal parts of $X$ to zero:
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \f1n \sum_{\ell>0} \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.
  \label{eq:error}
\end{align}
This reconstruction error can be maximized analytically, yielding a reconstruction error bound of the form
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2} \le \f{\epsilon_V^2}{n},
  \label{eq:bound}
\end{align}
where the complexity of computing the {\it quantum error scale} $\epsilon_V^2$ is the same as that of computing the classical error scale $\S_V^2$.
For comparison with the bound in Eq.~\eqref{eq:bound_eps}, we note that the error scale $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.

[\red{todo: finish discussion about $\epsilon_V^2$}]

As a final comment, we note that the quantum error scale only provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$, provided by Eq.~\eqref{eq:error}, may be considerably smaller.
While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho\to\tilde\rho_V$.

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V^2$ defined in Eq.~\eqref{eq:scale}.
If $\abs{V}$ is too large for such optimization  to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V^2$.
After choosing a set of measurement axes $V$, (experimentally) make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates of the expectation values $\bk{T_{\v v\ell,0}}_\rho$, which can be used to construct the state $\tilde\rho_V$ defined in Eq.~\eqref{eq:reconstructed_state}.
If $\tilde\rho_V$ has negative eigenvalues, its (Euclidean) distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.

@@

As a last point, we mention that the quantum error scale $\epsilon_V^2$ can also be used to evaluate the ``quality'' of measurement axes $V$ chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-error-scale sets of measurement axes with higher probability.
Two observations may explain the poor performance of these sparse signal recovery methods.
First, spin qudit tomography essentially boils down to recovering sparse signals with a particular harmonic structure, namely a fixed degree $\ell$.
The existing sparse signal recovery algorithms do not assume or account for any harmonic structure, and may therefore perform worse than expected on this particular subset of signals.
Second, spin qudit tomography requires recovering signals that are band-limited at degree $d$, but have $O(d)$ non-zero harmonic coefficients (out of $d^2$ coefficients in total).
The existing sparse signal recovery methods, on the other hand, may only perform well for signals with sparsity much smaller than $d$.
In any case, the fact that uniform sampling outperforms these methods only simplifies our randomized tomography protocol.

@@

which is block diagonalized as

@@

  U \equiv \sum_{\ell,m} \ket{\ell m} \obra{T_{\ell m}},

@@

  U M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},

@@

  \rho = \sum_{\ell,m} \bk{Q^V_{\ell m}}_\rho \p{Q^V_{\ell m}}^\dag,

@@

  \tilde\rho_V
  \equiv \sum_{\ell,m} \sp{\sum_i q^V_{\ell mi} \tilde T_{i\ell}}
  \p{Q^V_{\ell m}}^\dag
  \approx \sum_{\ell,m} \sp{\sum_i q^V_{\ell mi} \bk{T_{i\ell}}_\rho}
  \p{Q^V_{\ell m}}^\dag
  = \sum_{\ell,m} \bk{Q^V_{\ell m}}_\rho \p{Q^V_{\ell m}}^\dag

@@

We discuss the details of efficiently computing the error in Eq.~\eqref{eq:error_SM} in Appendix \ref{sec:compute}, but devote the rest of this Appendix to finding upper bounds for this error.

@@

  \sp{\bk{m|\N_{V\ell}|n} + \bk{n|\I\N_{V\ell}\I|m}}

@@

@@

  \oket{\rho_V^\star} = \sum_\ell \L_{V\ell}^{-1} \oket{\chi_{V\ell}},

@@

where $\L_{V\ell}^{-1}$ is the inverse of $\L_{V\ell}$.

@@

In order to establish simpler bounds on reconstruction error, we define the probability $p^i_\mu\equiv\bk{\Pi_{i\mu}}_\rho$, collect these probabilities into the classical probability distribution $p^i = \p{p^i_{-s}, p^i_{-s+1}, \cdots, p^i_s}$, and define the vector $t_\ell \equiv \p{t_{\ell,-s},t_{\ell,-s+1},\cdots,t_{\ell s}}$ of all $t_{\ell\mu}\equiv\bk{\mu|T_{\ell,0}|\mu}$.

@@

@@

  \equiv \sum_\mu p_\mu X_\mu^2 - \p{\sum_\mu p_\mu X_\mu}^2
is the weighted variance of $X$.

@@

which lets us bound

@@

@@

where $\S_V^2=\sum_\ell\S_{V\ell}^2$.

@@

  \le \f1n \sum_\ell \Gamma_\ell^2 \S_{V\ell}^2
  \le \f{\S_V^2}{2n},

@@

where $\S_V^2 = \sum_\ell \S_{V\ell}^2$.
The naive error bound in Eq.~\eqref{eq:bound_eps} of the main text therefore holds in the quantum case with the replacement $\epsilon^2\to1/2n$.

>>>>

\usepackage{dsfont} % for identity operator
\newcommand{\1}{\mathds{1}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Tomographic reconstruction error}
\label{sec:bound}

For the practically minded, the mere existence of solutions to a problem is less interesting than the exposition of a particular solution.
On a high level, a spin qudit tomography protocol consists of
\begin{enumerate*}
\item selecting a set of measurement axes,
\item collecting measurement data on spin projection onto these axes, and then
  \label{tomo:measure}
\item processing the collected data to reconstruct the state of the spin qudit.
\end{enumerate*}
Whereas step \ref{tomo:measure} can involve a host of platform-dependent technical challenges, in the following two sections we discuss the steps to take before and after collecting measurement data.

@@

A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\set{\Pi_{\v v\mu}}$ that will be measured in an experiment.
By a simple change of basis, measuring these projectors is equivalent to measuring the set of transition operators $\v T_V\equiv\set{T_{\v v\ell,0}}$.
By flattening the $d\times d$ matrices $T_{\v v\ell,0}$ in $\v T_V$ into $d^2$-component column vectors $\oket{T_{\v v\ell,0}}$, we can construct the {\it measurement matrix}
  \label{eq:meas_mat}
Here $\v v$ and $\ell$ label a column of $M_V$, or equivalently label a standard basis vector $\ket{\v v\ell}$ of a $\p{\abs{V}\times d}$-dimensional vector space.
A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured transition operators in $\v T_V$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
In this case $M_V$ must be full rank, with $d^2$ nonzero singular values.
Indexing these singular values $M^V_k$ and the corresponding (normalized) right singular vectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots,x^V_{k,\abs{V}})$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal qudit operators
  Q^V_k \equiv \sum_j q^V_{kj} T_j,
  \v q^V_k \equiv \f{\v x^V_k}{M^V_k},
where for shorthand we use a combined index $j=\p{\v v_j,\ell_j}$ to specify both a measurement axis $\v v_j$ and a degree $\ell_j$, which identify the transition operator $T_j\equiv T_{\v v_j\ell_j,0}$.
These operators allow us to expand any state $\rho$ of a $d$-level spin qudit in the form
  \rho = \sum_{k=1}^{d^2} \rho_k^V Q_k^V,
  \rho_k^V \equiv \bk{Q^V_k}_\rho.
Given empirical estimates $\tilde T_j$ of the expectation values $\bk{T_j}_\rho$, an empirical estimate $\tilde\rho_V$ of $\rho$ is then
  \tilde\rho_V \equiv \sum_k \tilde\rho^V_k Q^V_k,
  \label{eq:reconstructed_state}
  \tilde\rho^V_k \equiv \sum_j q^V_{kj} \tilde T_j
  \approx \sum_j q^V_{kj} \bk{T_j}_\rho
  = \bk{Q^V_k}_\rho
  = \rho^V_k.
The measurement matrix $M_V$ allows us to make concrete statements about the statistical error between the empirical estimate $\tilde\rho_V$ and the true state $\rho$.
Assume, for example, that the estimates $\tilde T_j$ are equal to $\bk{T_j}_\rho$ up to uncorrelated noise with variance at most $\epsilon^2$:
  \tilde T_j = \bk{T_j}_\rho + \epsilon_j,
  \bbk{\epsilon_j \epsilon_{j'}} \le \epsilon^2 \delta_{jj'},
  \label{eq:error_assumption}
where $\set{\epsilon_j}$ are independent random variables and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging.
In this case, the mean squared error with which $\tilde\rho^V_k$ approximates $\rho^V_k$ is
  \Bbk{\abs{\tilde\rho^V_k-\rho^V_k}^2}
  &= \Bbk{\p{\tilde\rho^V_k-\rho^V_k}^*\p{\tilde\rho^V_k-\rho^V_k}} \\
  &= \sum_{j,j'} \p{q^V_{kj}}^* q^V_{kj'}\,
  \bbk{\epsilon_j \epsilon_{j'}} \\
  &\le \sum_j \abs{q^V_{kj}}^2 \epsilon^2
  = \p{\f{\epsilon}{M^V_k}}^2.
Using the fact that the operators $Q^V_k$ are orthonormal, we can therefore bound the mean squared (Euclidean) distance between $\tilde\rho_V$ and $\rho$ as
  \le \epsilon^2 \S_V^2,
  \label{eq:bound_eps}
where $\norm{X} \equiv \sqrt{\obk{X|X}} = \sqrt{\tr\p{X^\dag X}}$ is the (Hilbert-Schmidt, or Frobenius) norm of $X$, and the {\it classical error scale}
  \S_V^2 \equiv \sum_k\f1{\p{M^V_k}^2} = \norm{M_V^{-1}}^2,
is the squared norm of the right inverse $M_V^{-1}$ of $M_V$, such that $M_V M_V^{-1}=1$.
Note that the classical error scale $\S_V^2$ diverges if $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.

Computing the classical error scale $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of a transition operator $T_{\ell m}$ is preserved under rotations, which implies that the unitary
\begin{align}
  U \equiv \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell
  \oket{T_{\ell m}} \bra{\ell m},
  \label{eq:trans_basis}
\end{align}
with vectors $\oket{T_{\ell m}}$ in a column indexed by integers $\p{\ell,m}$, block-diagonalizes the measurement matrix into $d$ blocks indexed by the degree $\ell$:
\begin{align}
  U^\dag M_V = \sum_{\ell=0}^{d-1} \op{\ell} \otimes M_{V\ell},
\end{align}
where the $\p{2\ell+1}\times\abs{V}$ blocks are
\begin{align}
  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell,0}} \bra{\v v}
  = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
\end{align}
As the singular values of $M_V$ are invariant under unitary transformations, it follows that
\begin{align}
  \S_V^2 = \sum_\ell \S_{V\ell}^2,
  &&
  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2,
\end{align}
where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time (see Appendix \ref{sec:compute} for details on efficiently computing $\S_{V\ell}^2$).
If we assume that $\abs{V}\sim d$, then computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).

@@

The assumption that observables can be estimated up to uncorrelated noise with maximal variance $\epsilon^2$, summarized by Eq.~\eqref{eq:error_assumption}, is reasonable when measurement error is dominated by classical sources of experimental noise.
However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:error}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$, with shot noise the dominant source of error.
In this case, the mean squared distance between the approximation $\tilde\rho_V$ and the true state $\rho$ can be compactly expressed using trace inner products of the {\it covariance matrices}
\begin{align}
  \C_\ell\sp{\rho} \equiv \sum_{m,n=-\ell}^\ell
  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n}
\end{align}
where $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$, and the {\it noise matrices}
\begin{align}
  \N_{V\ell} \equiv M_{V\ell} \diag\sp{M_{V\ell}^{-1}
    \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag,
\end{align}
where $\diag\sp{X}$ sets all off-diagonal parts of $X$ to zero:
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \f1n \sum_{\ell>0} \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.
  \label{eq:error}
\end{align}
Maximizing this error over all $\rho$ (for a fixed set of axes $V$) yields a reconstruction error bound of the form
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2} \le \f{\epsilon_V^2}{n},
  \label{eq:bound}
\end{align}
where the complexity of computing the {\it quantum error scale} $\epsilon_V^2$ (defined below) is the same as that of computing the classical error scale $\S_V^2$ (see Figure \ref{fig:times}).
For comparison with the bound in Eq.~\eqref{eq:bound_eps}, we note that $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
We also note that the (squared) reconstruction error in Eq.~\eqref{eq:error} and its bound in Eq.~\eqref{eq:bound} obey the ``standard quantum limit'' of $\sim1/\sqrt{n}$ scaling in the number of measurements.
In principle, this scaling can be improved to $\sim1/n$ by preparing and measuring entangled copies of many qudits \cite{giovannetti2006quantum}.

The quantum error scale $\epsilon_V^2$ only provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$, provided by Eq.~\eqref{eq:error}, may be considerably smaller.
While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho\to\tilde\rho_V$.

The remainder of this section is devoted to defining $\epsilon_V^2$, and can be skipped without consequence for the discussions in the rest of this paper.
To write out the quantum error scale $\epsilon_V^2$ explicitly, we define a matrix of structure constants
\begin{align}
  g_{L\ell} \equiv \sum_{m,n} g^L_{\ell mn} \op{m}{n},
  &&
  g^L_{\ell mn} \equiv \obk{T_{L,m+n}|T_{\ell m}T_{\ell n}},
\end{align}
and the inversion operator $\I\equiv\sum_m\p{-1}^m\op{m}{m}$, which define components of the $d^2$-component vector
\begin{align}
  \ket{\chi_V}
  &\equiv \sum_{L=0}^{d-1} \sum_{M=-L}^L \chi^V_{LM} \ket{LM}, \\
  \chi^V_{LM} &\equiv \sum_\ell \obk{\N_{V\ell}|\D_M|\I g_{L\ell}}, \\
  \D_M &\equiv \sum_{m,n} \delta_{M,m+n} \op{mn}.
\end{align}
We then define the state vector
\begin{align}
  \oket{\rho_V^\star} = \f1d \oket{\1_d} + U \sum_{\ell>0} \sp{\N_{V\ell} + \I\N_{V\ell}^*\I}^{-1} \ket{\chi_{V\ell}},
\end{align}
where $\1_d$ is the $d\times d$ identity matrix, $\ket{\chi_{V\ell}} \equiv \sum_{m=-\ell}^\ell \ket{\ell m} \bk{\ell m|\chi_V}$ is the projection of $\ket{\chi_V}$ onto the subspace of fixed degree $\ell$, the unitary $U$ defined in Eq.~\eqref{eq:trans_basis} takes $\ket{\ell m}\to \oket{T_{\ell m}}$, and $\N_{V\ell}^*$ is the complex conjugate of $\N_{V\ell}$.
The quantum error scale $\epsilon_V^2$ is then
\begin{align}
  \epsilon_V^2 \equiv \sum_{\ell>0} \obk{\N_{V\ell}|\C_\ell\sp{\rho_V^\star}}.
  \label{eq:scale}
\end{align}
We derive Eq.~\eqref{eq:scale} in Appendix \ref{sec:error}, and discuss its efficient calculation in Appendix \ref{sec:compute}.
Note that the ``state'' $\rho_V^\star$, which maximizes the reconstruction error in Eq.~\eqref{eq:error}, has trace $\tr\rho_V^\star=1$, but may generally have negative eigenvalues.
In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing reconstruction error over the (restricted) domain of states $\rho$ with a non-negative spectrum.

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V^2$ defined in Eq.~\eqref{eq:scale}.
If $\abs{V}$ is too large for such optimization  to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V^2$.
After choosing a set of measurement axes $V$, (experimentally) make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates of the expectation values $\bk{T_{\v v\ell,0}}_\rho$, which can be used to construct an estimate $\tilde\rho_V$ of $\rho$, defined in Eq.~\eqref{eq:reconstructed_state}.
If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.

@@

As a last point, we mention that the quantum error scale $\epsilon_V^2$ can also be used to evaluate the ``quality'' of measurement axes $V$ chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-error-scale sets of measurement axes with higher probability [\red{todo: add figure}].
Two observations may explain the poor performance of these sparse signal recovery methods.
First, spin qudit tomography essentially boils down to recovering sparse signals with a particular harmonic structure, namely a fixed degree $\ell$.
The existing sparse signal recovery algorithms do not assume or account for any harmonic structure, and may therefore perform worse than expected on this particular subset of signals.
Second, spin qudit tomography requires recovering signals that are band-limited at degree $d$, but have $O(d)$ non-zero harmonic coefficients (out of $d^2$ coefficients in total).
The existing sparse signal recovery methods, on the other hand, may only perform well for signals with sparsity much smaller than $d$.
In any case, the fact that uniform sampling outperforms these methods only simplifies our randomized tomography protocol.

@@

which can be block diagonalized as

@@

  U^\dag M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},

@@

  U \equiv \sum_{\ell,m} \oket{T_{\ell m}} \bra{\ell m},

@@

  \rho = \sum_{\ell,m} \bk{Q^V_{\ell m}}_\rho Q^V_{\ell m},

@@

  \tilde\rho_V \equiv \sum_{\ell,m}
  \sp{\sum_i q^V_{\ell mi} \tilde T_{i\ell}} Q^V_{\ell m}
  \approx \sum_{\ell,m}
  \sp{\sum_i q^V_{\ell mi} \bk{T_{i\ell}}_\rho} Q^V_{\ell m}
  = \sum_{\ell,m} \bk{Q^V_{\ell m}}_\rho Q^V_{\ell m}

@@

We discuss the details of efficiently computing the error in Eq.~\eqref{eq:error_SM} in Appendix \ref{sec:compute}, and devote the rest of this Appendix to finding upper bounds for this error.

@@

  \bk{m|\p{\N_{V\ell} + \I \N_{V\ell}^* \I}|n}

@@

Here $\N_{V\ell}^*$ is the complex conjugate (and also the transpose) of $\N_{V\ell}$.

@@

  \oket{\rho_V^\star}
  = \f1d \oket{\1_d} + \sum_{\ell>0} \L_{V\ell}^{-1} \oket{\chi_{V\ell}},

@@

where $\1_d$ is the $d\times d$ identity matrix (added artificially to enforce $\tr\rho_V^\star=1$; the identity component of $\rho_V^\star$ does not actually contribute to reconstruction error), and $\L_{V\ell}^{-1}$ is the inverse of $\L_{V\ell}$.

@@

In order to find a simple bound on reconstruction error, we define the probability $p^i_\mu\equiv\bk{\Pi_{i\mu}}_\rho$, collect these probabilities into the classical probability distribution $p^i = \p{p^i_{-s}, p^i_{-s+1}, \cdots, p^i_s}$, and define the vector $t_\ell \equiv \p{t_{\ell,-s},t_{\ell,-s+1},\cdots,t_{\ell s}}$ of all $t_{\ell\mu}\equiv\bk{\mu|T_{\ell,0}|\mu}$.

@@

  \equiv \sum_\mu p_\mu X_\mu^2 - \p{\sum_\mu p_\mu X_\mu}^2,

@@

where $\sigma_p^2\p{X}$ is the weighted variance of $X$.

@@

and lets us bound

@@

  \t{and}
  \Bbk{\norm{\tilde\rho_V-\rho}^2} \le \f{\S_V^2}{n}.

@@

where $\S_V^2=\sum_\ell\S_{V\ell}^2=\norm{M_V^{-1}}^2$.

@@

  \le \f{\S_V^2}{2n}.

@@

The naive error bound in Eq.~\eqref{eq:bound_eps} of the main text therefore still holds with the replacement $\epsilon^2\to1/2n$.
